\documentclass[parskip=full]{uvamscse}

\setcounter{secnumdepth}{4}

\titleformat{\paragraph}
{\normalfont\normalsize\bfseries}{\theparagraph}{1em}{}
\titlespacing*{\paragraph}
{0pt}{3.25ex plus 1ex minus .2ex}{1.5ex plus .2ex}

\input{program-listings}
\newcommand{\cmd}[1]{\texttt{$\backslash$#1}}

\title{Detecting Violations of CSS Code Conventions}
% \coverpic[100pt]{figures/terminal.png}
% \subtitle{}
% \date{Spring 2014}


\author{Boryana Goncharenko}
\authemail{boryana.goncharenko@gmail.com}
% \host{Grammarware, Inc., \url{http://grammarware.github.io}}

\abstract{    

Code conventions are used to preserve code base consistency and express preference of a particular
programming style. Often, code conventions are expressed in natural language and it is a
responsibility of the developers to read, understand and apply them. Typically, developers need to
ensure that their code complies to a given style guide manually. There are a number of tools that
can automatically detect violations of conventions, however, current solutions remain rigid,
laborious or with limited scope.

This thesis aims at answering three research questions. First, it evaluates the need for CSS
conventions based on whether CSS is still handcrafted. Second, it discovers existing code
conventions. Third, it designs a domain-specific language that is capable of expressing existing CSS
code conventions. The thesis presents a specification of the designed domain-specific language and an implementation of its interpreter that detects violations automatically.

}


\begin{document}
\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{ch-intro}
\input{ch-back}
\input{ch-eval}
\input{ch-discover}
\input{ch-express}
\input{ch-conclusion}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


{%\tiny
\bibliographystyle{ieeetr}
\bibliography{thesis}
}

\end{document}
